---
title: "Wednesday R markdown"
author: "Mitch Korolev"
date: "2023-03-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Libraries

```{r libraries}
library(dplyr)
library(tidyr)
library(readr)
```

## Load our data!

```{r dataset}

catch.original <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1")
```

The select command from the dplyr package selects **columns**.

The filter command from the dplyr package filters **rows**.

## Clean data

```{r cleaning data}
#Remove the All and notesRegCode columsn from the catch_original data frame

catch_data <- catch.original %>% 
    select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum)

# catch_data <- catch_original %>% 
#   select(-All, -notesRegCode)

head(catch_data)

```

```{r}
#Mutate chinook class from character to numeric
catch_clean <- catch_data %>% 
  mutate(Chinook = as.numeric(Chinook))

#search for the NA value
i <- which(is.na(catch_clean$Chinook))

# located our NA value! It's a uppercase I
catch_data[i, ]

catch_clean <- catch_data %>%
  mutate(Chinook = if_else(Chinook == "I", "1", Chinook)) %>% 
  mutate(Chinook = as.integer(Chinook))
```

## change shape of data using pivot longer and pivot wider

```{r}
#pivot longer practice

catch_long <- catch_clean %>% 
  pivot_longer(cols = -c(Region, Year), names_to = 'species', values_to = 'catch')

View(catch_long)

# pivot wider practice

catch_wide <- catch_long %>% pivot_wider(names_from = species, values_from = catch)

head(catch_wide)
```


## Practice renaming columns
```{r}
catch_long <- catch_long %>% 
  rename(catch_thousands = catch)

head(catch_long)

# remove catch_thousands

catch_long <- catch_long %>% 
  mutate(catch = catch_thousands * 1000) %>%
  select(-catch_thousands)

head(catch_long)
```

## practice using group_by and summarize.

```{r}
mean_region <- catch_long %>%
  group_by(Region) %>%
  summarize(catch_mean = mean(catch))

head(mean_region)

# summarize count by region

n_region <- catch_long %>%
  group_by(Region) %>%
  summarize(number_of_observations =n(), 
            total_catch = sum(catch))

head(n_region)
```

## Lesson on filtering
```{r}
ALU_catch <- catch_long %>% filter(Region == "ALU")

head(ALU_catch)
```

## Arrange some rows! sad-woo

```{r}
mean_region_arrange <- catch_long %>% 
  group_by(Region) %>% 
  summarize(mean_catch = mean(catch)) %>%
  arrange(mean_catch)

head(mean_region_arrange)
```



## We're practicing our joins

```{r}
# loading in auxillary dataset
region_defs <- read.csv('https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1')

head(region_defs)

```

##Join our data

```{r}
catch_joined <- left_join(catch_long, region_defs, by = c("Region" = "code"))

head(catch_joined)
```

```{r}
region_defs <- region_defs %>%
  rename(Region = code, Region_Name = mgmtArea)

catch_joined <- left_join(catch_long, region_defs, by = "Region")

head(catch_joined)
```


```{r}
sites_df <- data.frame(site = c("HAW-101",
                                "HAW-103",
                                "OAH-320",
                                "OAH-219",
                                "MAI-039"))

sites_df %>% 
```


